Online Verification of Commutativity

### 2020-06-30 acmPaper/paper.tex
cleanup i think

We start with the set of all parallel pairs in the diagram (with the new edge added in), and pare it down to be finite by handling cycles.
We find all simple cycles in the diagram, using a procedure like Johnson's algorithm. 
%TODO cite the paper.
We verify for each cycle that a single traversal starting on any node is equal to the identity function by adding pairs of (cycle starting at node, identity function) for each node in the function to the output verification set.
\xxx[AK]{Need cleaner way to put this.}
As cycles must now be the identity, for any pair in the set of all parallel paths, any instance of a cycle can be removed to obtain an equivalent pair with shorter, cycle free paths.
If the shorter pair has equal paths then the paths in the original pair must also be equal to each other.
It is therefore safe to remove all pairs of paths with cycles, leaving only parallel pairs where neither path has a cycle.
Since there are only a finite number of paths in the graph with no cycles, (to the order of n! as a path is an ordering on nodes), and the parallel pairs are a subset of the possibilities generated by choosing a first and second member from this finite set, the resultant set of all parallel pairs must be finite.
%TODO Check if O(n!n) = O(n). What about O((n!)^2)?
The output of the algorithm, then, is the union of the set of cycle verification pairs, and the parallel pairs found by doing an all-acyclic-path search from every node in the diagram with the new edge added.

@@

\paragraph{Implementation}
\xxx[AK]{Copy in python code from implementation? Appendix?}
\xxx[DG]{Not sure we need this paragraph, the implementation is pretty clear from the description of the algorithm}

This algorithm is also in the worst case as bad as factorial in the number of nodes. However, in the average case, it significantly outperforms the naive batch baseline. For sparse diagrams it is even competitive with our ultimate polynomial solution. Average case heuristic analysis will be presented in the evaluation section.

\subsection{Optimal batch solution}
Murota's main result ~\cite{commutative} is a solution to the question of whether a given acyclic diagram commutes that returns the minimal number of equality checks. The paper describes the following algorithm to find the ($V^2E$ bounded) minimal set of pairs that needs to be checked.

@@

For bilinking $s$, let $F(s)$ be the vector in $\mathbb{F}2^{|E|}$ representing the edges present in s (the $n^{th}$ dimension of $F(s)$ is 1 if the corresponding edge is in $s$, and 0 otherwise). Let this function be extended to sets, so that for some set of bilinkings $S, F(S) = \{ F(s) | s\in S \}$.

for a set of bilinkings $r$, the closure function $cl$ is defined as:
\xxx[dg]{I think this definition is incomplete}
$cl(r) = \{ s\in R_0| s$ is linearly dependent on $F(r) \}$.

>>>>

We start with the set of all parallel pairs in the diagram (with the new edge added in), and pare it down to be finite by handling cycles.
We find all simple cycles in the diagram, using a procedure like Johnson's algorithm. 
\xxx[AK]{TODO cite the paper.}
We verify for each cycle that a single traversal starting on any node is equal to the identity function by adding pairs of (cycle starting at node, identity function) for each node in the function to the output verification set.
\xxx[AK]{Need cleaner way to put this.}
As cycles must now be the identity, for any pair in the set of all parallel paths, any instance of a cycle can be removed to obtain an equivalent pair with shorter, cycle free paths.
If the shorter pair has equal paths then the paths in the original pair must also be equal to each other.
It is therefore safe to remove all pairs of paths with cycles, leaving only parallel pairs where neither path has a cycle.
Since there are only a finite number of paths in the graph with no cycles, (to the order of n! as a path is an ordering on nodes), and the parallel pairs are a subset of the possibilities generated by choosing a first and second member from this finite set, the resultant set of all parallel pairs must be finite.
\xxx[AK]{TODO Check if O(n!n) = O(n). What about $O((n!)^2)$?}
The output of the algorithm, then, is the union of the set of cycle verification pairs, and the parallel pairs found by doing an all-acyclic-path search from every node in the diagram with the new edge added.

@@

\paragraph{Analysis}
This algorithm is also in the worst case as bad as factorial in the number of nodes. However, in the average case, it significantly outperforms the naive batch baseline. For sparse diagrams it is even competitive with our ultimate polynomial solution. Average case heuristic analysis will be presented in the evaluation section.

\subsection{Optimal batch solution}
Murota's main result ~\cite{commutative} is a solution to a batch version of \textsc{verification set} - given an acyclic diagram, it returns the minimal set of equality checks that would succeed if and only if the diagram commutes.
The paper describes the following algorithm to find the ($V^2E$ bounded) minimal set of pairs that needs to be checked.

@@

For bilinking $s$, let $F(s)$ be the vector in $\mathbb{F}2^{|E|}$ representing the edges present in s (the $n^{th}$ dimension of $F(s)$ is 1 if the corresponding edge is in $s$, and 0 otherwise).
Let this function be extended to sets, so that for some set of bilinkings $S, F(S) = \{ F(s) | s\in S \}$. A notion of linear independence in this vector field exists.

For a set of bilinkings $r$, the closure function $cl$ is defined as:
\xxx[dg]{I think this definition is incomplete}\xxx[ak]{sorry, what is it missing?}
$cl(r) = \{ s\in R_0| s$ is linearly dependent on $F(r) \}$.
